---
title: '8001 Lab 1: Importing and Summarizing Data'
author: "Ben Burnley"
date: "2022-09-01"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Welcome to our lab section for Quantitative Methods! A few helpful tips before we get going. The box right below is called a `chunk` of code. Create a new one using the green +C box at the top right or using the shortcut "option + command + i" on Mac or "control + alt + i" on Windows.

```{r}
### this is a new code chunk! 
### use hashtags to add comments on your code 
### tip one is to always leave comments on your code
```

A couple things...

-   Code in R is case-sensitive, so be mindful of your naming conventions!

-   Run lines of code by holding down "ctrl" + "r" OR "ctrl" + "enter" on Windows; "cmd" + "return" on Mac

-   There is also a button up top that says "Run" and you can use that too!

-   RMarkdown is a great way to create high quality documents with text, data analysis, and data visualization capabilities all in one place.

-   In RMarkdown, run code `chunks` by clicking the green play button at the top right of the `chunk` when you hit the blue yarn `knit` button at the top left of the file, all r code will be executed as part of generating your output file.

## Creating Simple Objects

### Scalars (single value objects):

```{r}
x <- 5 #creating object "x" with value of 5
y <- 3 #creating object "y" with value of 3

x #tells us value of x is 5
y #tells us value of y is 3
```

### Creating Vectors (objects that are sequences of values):

```{r}
a <- c(2, 4, 6, 8) #use c to combine values into vector
b <- c("cat", "dog", "bird", "fish") #creating a character vector
```

### Creating Dataframes:

```{r}
age <- c(2, 4, 6, 8) #use c to combine values into vector
animal<- c("cat", "dog", "bird", "fish") #creating a character vector

data.frame(animal,age) # create the dataframe
my_pets <- data.frame(animal,age) #store the dataframe
```

### Creating Factors

```{r}
# normal character vector
classes <- c("freshman", "sophomore", "junior", "senior")
classes

# create factor variable 
classes_ordinal <- factor(classes, levels = c("senior", "junior", "sophomore", "freshman"))
classes_ordinal
```

## Datasets We Are Using Today

-   Ideology score of U.S. legislators for the 117th Congress

    -   `HS117_members.csv`
    -   <https://voteview.com/data>

-   Ideology score of countries using United Nations General Assembly votes

    -   `IdealpointestimatesAll_Mar2021.tab`
    -   <https://dataverse.harvard.edu/dataset.xhtml?persistentId=hdl:1902.1/12379>

## Let's Describe our data

-   For code chunks with data import, we typically set `echo = FALSE`, because data import doesn't convey anything that needs displaying. Dataset import will also generate a lot of annoying messages that will be displayed in your pdf unless you set `message = FALSE` in the code chunk as well
-   For problem sets, I am not interested in how you are importing data UNLESS you are having a problem with data import.

### Importing Data

```{r, echo = FALSE, message = FALSE}
### SET WORKING DIRECTORY 
##
getwd()
setwd("~/Desktop/8001")
#### IMPORTING THE DATASET
##
#install.packages(readr) #remember, only once/installation of R
library(readr)
# Read .csv file
# Obtain the csv file from: `HS117_members.csv` at  https://voteview.com/data

voteview <- read_csv("HS117_members.csv") # load in voteview data
UNideal <- read_csv("IdealpointestimatesAll_Jun2022.csv") #load in UN data
```


```{r, echo = FALSE, message = FALSE, eval = FALSE}
### OTHER EXAMPLES (code will not run, this is for your reference)
### Excel Files
#install.packages("readxl")
library(readxl)

# Importing an Excel file with only one sheet
ExampleExcel <- read_excel("ExampleExcel.xlsx")

# Importing one sheet from an Excel file with multiple sheets
ExampleExcelSheet <- read_excel("ExampleExcel.xlsx", sheet = "Sheet Name")


#install.packages("haven")
library("haven")
# If you are having trouble installing "haven" you can try downloading "tidyverse" (big package that includes "haven")

# Stata files
ExampleStata <- read_dta("ExampleStata.dta")

# SAS files
ExampleSAS <- read_sas("ExampleSAS.sas7bdat") 

# SPSS files
ExampleSPSS <- read_sav("ExampleSPSS.sav") 

## R files

load(file = "ExampleR.RData")

# two more for last resort
read_table()
read_delim()
```

## Desribing the dataset:

To see a list of all of the variable names in the dataset (object) called "voteview":

```{r}
names(voteview)
```

To look at a spreadsheet version of your dataset (it will open in new tab):

```{r, eval=FALSE}
View(voteview) # note the capital V
```

Note the `eval = FALSE` which instructs R not to try and run this line of code in `pdf` generation. Running the `chunk` with the green button will still work.

To see a snapshot of your data:

```{r}
head(UNideal) # shows you the first 6 rows
```

Understanding check: what is our unit of observation in this dataset?

To see dimensions of dataset:

```{r}
dim(UNideal)
```

To see number of rows:

```{r}
nrow(UNideal)
```

To see number of columns:

```{r}
ncol(UNideal)
```

Understanding check: How many variables does the dataset have? How many observations?

## Accessing Variables:

There are two approaches to calling up specific variables from your dataset: 1) Using "attach" function 2) Specifying location of object

1)  The "attach" function seems easier, but can create problems when using multiple datasets. See <https://www.r-bloggers.com/to-attach-or-not-attach-that-is-the-question/>

```{r}
attach(voteview)
summary(nominate_dim1)
detach(voteview)
```

2)  Rather than using "attach" function, it is usually better to specify object's location. To specify a variable within a dataset, use `$`.

```{r}
summary(voteview$nominate_dim1)
```

## Describing Variables:

Checking variable types

```{r}
class(UNideal$session) # example of numeric variable
class(UNideal$Countryname) # example of character variable
```

## Describing numeric variables

First, I'm going to get rid of some annoying data so that I don't have to include `na.rm = TRUE` in every line

```{r}
voteview <- voteview[-229,]
```

What did this code do?

```{r}

summary(voteview$nominate_dim1) # 5-number summary (min, 1st, median, 3rd, max)

sum(is.na(voteview$nominate_dim1)) # counting number of missing values

mean(voteview$nominate_dim1) # mean value

min(voteview$nominate_dim1) # minimum value

max(voteview$nominate_dim1) # maximum value

range(voteview$nominate_dim1) # minimum and maximum

var(voteview$nominate_dim1) # variance

sd(voteview$nominate_dim1) # standard deviation
  sqrt(var(voteview$nominate_dim1)) # (also standard deviation)

IQR(voteview$nominate_dim1) # interquartile range (3rd-1st quartile)

```

-   Note that base R does not have a function for finding the mode.
-   To do this easily, we can use the "modeest" package.

Only run this chunk once:

```{r, eval=FALSE}
install.packages("modeest") # installing modeest package once/R installation
```

Then you can use this function:

```{r}
library("modeest") # activating modeest package
mfv(voteview$nominate_dim1) # mfv = "most frequent value"
```

## Other ways to describe variables

```{r}
unique(voteview$state_abbrev) # lists unique values for a variable

length(unique(voteview$state_abbrev)) # counts number of unique values

table(voteview$state_abbrev) # frequency table for the number of observations for each value of region

table(voteview$nominate_dim1) # not always useful for examining numeric variables!
```

Exercises

Exercise 1

```{r}
vector1 <- c()
vector2 <- c()
vector3 <- c()

mydata <- data.frame(vector1, vector2, vector3)
```
Exercise 2 

Compare the mean and the standard deviation of `nominate_dim2` variable to the mean and standard deviation for `nominate_dim1`. Find the mean and standard deviation of nominate_dim2 for only House members and for only Senate members.

```{r}
```
