---
title: "Lab 2: Exploratory Data Analysis"
author: "Ben Burnley"
date: "2022-09-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Review from Week 1

We will use the mtcars data set for the examples below, so let's go ahead and `attach` it.

```{r}
attach(mtcars)
help(mtcars)

#use the data set mtcars for this analysis. how can we see the first part of this dataset?


# what are the variable names in this dataset?


# get summary statistics for the "disp" variable?


# how do we find the variance of the "wt" variable?


# how many values are possible for the variable "cyl"?


# what is "cyl" a factor? 

```

Ok...

On to this week's material!

# Exploratory Data Analysis

## Bivariate Descriptive Statistics

Last week, we looked at ways univariate descriptive statistics. Let's add in the bivariate descriptive statistics we covered this week in class. We will continue using the `mtcars` data set.

```{r}
# find the covariance of two variables 
cov(wt, disp)

# find the correlation between the same two variables
cor(wt, disp)

# a more under the hood way of understanding correlation
cov(wt, disp)/(sd(wt)*sd(disp))

# what about a variance/covariance matrix?
cov(mtcars)

# comprehension check: the matrix shows variance on the diagonal, and covariance in remaining cells. The first two values in the mpg column are:
var(mpg)
cov(mpg, cyl)
```

## Exploratory Visualization

Looking at summary statistics is helpful, but can obscure underlying trends in the data. Visualizing helps us to better understand the data and the patterns (or lack of patterns) present.

### Univariate Plots

```{r}
# histograms
hist(mpg)

# working with breaks
hist(mpg, breaks = 2)
hist(mpg, breaks = 25)

# add some color!
hist(mpg, breaks = 10, col = "blue")

```

```{r}
# density plot 
plot(density(mpg), main = "Density of `mpg` variable")

#NOTE: this is not ideal for a sample of 32. Becomes helpful with 1) large sample sizes and 2) when comparing distributions/densities. 
```

```{r}
# box plots 
boxplot(mpg)

# display horizontally
boxplot(mpg, horizontal = TRUE)

# add title
boxplot(x = mpg, horizontal = TRUE, main = "Boxplot of Miles Per Gallon")

# label x-axis
boxplot(x = mpg, horizontal = TRUE, main = "Boxplot of `mpg` Variable", xlab = "Miles Per Gallon")

# add color, this time with a hex code 
boxplot(x = mpg, horizontal = TRUE, main = "Boxplot of `mpg` Variable", xlab = "Miles Per Gallon", col = "#FF0000")
```

### Bivariate Plots

```{r}
plot(wt, disp)

# editing the x and y axes
plot(wt, disp, xlim = c(0,6), ylim = c(0,500))

# make it look nice
plot(wt, disp, frame.plot = F, pch = 16, col = "red", main = "Displacment vs. Weight from `mtcars`", xlab = "Weight (1000 lbs)", ylab = "Displacement (cu.in.)")
```

#### Exercises

```{r}
# create a histogram for a variable we haven't used

# create a boxplot for a variable we haven't used

# create a scatter plot for variables we haven't used

```

## Generating Data

It can be incredibly helpful to understand how to generate your own data and simulate data generation processes.

```{r}
detach(mtcars)

# for additional reading on the the functions we will use 
help(sample)
help(rnorm)
help(rbinom)
help(runif)

# Let's simulate a vote in the U.S. House of Representatives

# First, we need to set our seed so that we all get the same results 
set.seed(8001)

# create categories via sampling 
parties = c("Republican", "Democrat")
party = sample(parties, replace = TRUE, size = 435)

# create a approval variable using rnorm
approval = rnorm(435, mean = 50, sd = 10) # NOTE: default for rnorm is mean = 0, sd = 1

# create a roll call variable
vote = rbinom(435, size = 1, prob = .5)

# number of children 
kids = round(runif(435, 0, 5))

# put our variables together into a dataframe
data.frame(party, approval, vote, kids)
```

```{r}
#summary of party
table(party)

# summary of approval
hist(approval)
summary(approval) 

# summary of vote
hist(vote) # not so useful
table(vote)

# summary of kids 
summary(kids)
boxplot(kids, horizontal = T)
```

```{r}
# is there a relationship between kids and approval? 
plot(kids, approval)
boxplot(approval ~ kids, horizontal = T)
```
```{r}
# what if we wanted to create a relationship? 
funds = approval * 1000 + rnorm(435, mean = 10000, sd = 10000)

plot(approval, funds)

cov(approval,funds)
cor(approval,funds)
```

