---
title: "lab_5_regression_1"
author: "Ben Burnley"
date: "2022-09-29"
output: html_document
---

```{r}
# packages 
library(tidyverse)
library(readxl)

# load data 
pres_vote <- read_excel("pres_vote.xlsx")
ev <- read_csv(file = "econ_vote.csv")
```

## A Quick Tutorial on Distributions

You will need to know how to use these commands from time to time. The two you will use most are `pnorm` and `rnorm`, but all four come in handy. You can use the same commands for other distributions (`?distributions`)

`dnorm`: density function of the normal distribution

`pnorm`: cumulative density function of the normal distribution

`qnorm`: quantile function of the normal distribution

`rnorm`: random sampling from the normal distribution

### `pnorm` and `rnorm`

```{r}
# for help 
?distributions

# using pnorm
# for when you know a value and you want a probability of it occurring 
pnorm(1.96, mean = 0, sd = 1, lower.tail = TRUE) # mean = 0, sd = 1 are defaults, written here for clarity 
pnorm(1.96, mean = 0, sd = 1, lower.tail = FALSE)

# simulation using r norm 
data = rnorm(1000, mean = 50, sd = 10)
hist(data)

# in what percent of cases are observations greater than 80?
pnorm(80, mean = 50, sd = 10, lower.tail = F)

# in what percent of cases are observations lower than 30?
pnorm(30, mean = 50, sd = 10)

# in what percent of cases are observations between 45 and 55?
pnorm(55, 50, 10) - pnorm(45, 50, 10)

# in what percent of cases are observations above 55 AND below 45?


```

### `qnorm`

```{r}
# using qnorm: you have a quantile and you want a value
summary(data)
```

```{r}
# some qnorm examples 
qnorm(.25, 50, 10)
qnorm(.75, 50, 10)
qnorm(.25, 50, 10, lower.tail = F)
qnorm(.50, 50, 10)


# another way to think about this is you are giving R a probability

# what is the value such that 2.275013% of cases are less than it?
qnorm(0.02275013, 50, 10)
```

### `dnorm`

```{r}
# how often did a value occur? 
dnorm(50, 50, 10)

df = data.frame(data)
ggplot(df, aes(x = data))+
  geom_density()

data2 = rnorm(1000000)
df2 = data.frame(data2)
ggplot(df2, aes(x = data2))+
  geom_density()
dnorm(0)

```

## Regression

```{r}
## "Regress y on x"
?lm
# General command for OLS
regression1 <- lm(data$y ~ data$x) 
# An easier way to write the same thing:
regression1 <- lm(y ~ x, data = data) 

# View summary of regression results 
summary(regression1)
```

```{r}
# view our data set
View(pres_vote)

# Describe what is happening in this regression?
model1 <- lm(Trump_2016 ~ rnormvote, data = pres_vote)

# what is an lm object? 
class(model1)

View(model1)

summary(model1)
```

```{r}
# Fitted or predicted values are where a given x-value falls on the line of best fit
# Find values by substituting a given value of x into the regression equation
# pulling out fitted values - both commands work 
model1$fitted.values
fitted(model1)

# Residuals are the difference between the observed data and the fitted values
# Positive values mean prediction was too low, negative values mean prediction was too high
# pulling out residuals from a model objects - both commands work
model1$residuals
residuals(model1)


# plot residuals vs. fitted values 
# Plot
yhat <- model1$fitted.values
e <- model1$residuals 

plot(e ~ yhat) # can add graphing elements like title, etc.!
abline(h=0,col="red") # points on the line were perfectly predicted by our model

# in ggplot 
ggplot()+
  geom_point(aes(x = yhat, y = e))+
  geom_hline(yintercept = 0, color = "red")

# compare fitted values to observed values
ggplot()+
  geom_point(aes(x = yhat, y = pres_vote$Trump_2016))+
  geom_abline(color = "red")
```

```{r}
# an example of how looking at this might be helpful
x = seq(1:100) 
y = x^2
plot(x,y)
ex_model = lm(y ~ x)
plot(fitted(ex_model), residuals(ex_model))
abline(h = 0, col = "red")
```

## Example from Regression Lecture

```{r}
# Example from Regression lecture
View(ev)
range(ev$elec_margin, na.rm = TRUE)
range(ev$gdp_growth, na.rm = TRUE)

# Plot X against Y
plot(ev$gdp_growth, ev$elec_margin, xlim = c(-3, 3), ylim = c(-25, 25))
```

```{r}
# Regress election margin on gdp growth 
ev_fit <- lm(formula = elec_margin ~ gdp_growth, data = ev)


summary(ev_fit)
```

```{r}
# clean up our data 
ev_gg = ev %>% 
  filter(year != 2020) %>% 
  mutate(predicted = predict(ev_fit), # predict function uses model to create fitted values
         residuals = residuals(ev_fit)) 

# plot of linear regression model 
ggplot(ev_gg, aes(gdp_growth, elec_margin))+
  geom_point()+
  geom_smooth(method = "lm", se = F, color = "red")+
  theme_bw()
```

```{r}
ggplot(ev_gg, aes(x = gdp_growth, y = elec_margin)) +
  geom_smooth(method = "lm", se = FALSE, color = "lightgrey") +  # Plot regression slope
  geom_segment(aes(xend = gdp_growth, yend = predicted), alpha = .2) + # Segments show residual distance
  geom_label(aes(x= gdp_growth, y=(predicted+elec_margin)/2, label = round(residuals,3)), size = 3)  + # Labels show residual value                                                                  # alpha to fade lines
  geom_point() + # Values  
  geom_point(aes(y = predicted), shape = 1) + # Fitted values
  theme_bw()
```

```{r}
# sum of the squared residuals
sum((ev_gg$residuals)^2)
```

```{r}
## What if we used a different line? 

ev_gg = ev_gg %>% 
  mutate(predict2 = -.9292 + gdp_growth*10, #gdp growth increase margin by 10 points
         residual2 = elec_margin - predict2,
         predict3 =  -.9292 + gdp_growth*3, # gdp growth increases margin by 3 points
         residual3 = elec_margin - predict3)

```

```{r}
# The same figure with our first new line
ggplot(ev_gg, aes(x = gdp_growth, y = elec_margin)) +
  geom_smooth(method = "lm", se = F, color = "lightgrey")+
  geom_abline(intercept = -.9292, slope = 10, color = "red") +  # Plot regression slope
  geom_segment(aes(xend = gdp_growth, yend = predict2), alpha = .2) +  # Segments show distance of values from fitted values, alpha fades lines
  geom_label(aes(x= gdp_growth, y=(predicted+elec_margin)/2, label = round(residual2,3)), size = 3)  + # Labels show residual value                                                                  # alpha to fade lines
  geom_point() +   # Values                                                    
  geom_point(aes(y = predict2), shape = 1) +# Fitted values
  theme_bw()  # Add theme for cleaner look
```

```{r}
# sum of the squared residuals - linear model 
sum((ev_gg$residuals)^2)
# sum of the squared residuals - our guess of 10
sum((ev_gg$residual2)^2)
```

```{r}
ggplot(ev_gg, aes(x = gdp_growth, y = elec_margin))+
  geom_smooth(method = "lm", se = F, color = "lightgrey")+
  geom_abline(intercept = -.9292, slope = 3, color = "red") +  # Plot regression slope
  geom_segment(aes(xend = gdp_growth, yend = predict3), alpha = .2) +  # Segments show distance of values from fitted values 
  # Alpha to fade lines
  geom_label(aes(x= gdp_growth, y=(predicted+elec_margin)/2, label = round(residual3,3)), size = 3)  + # Labels show residual value              
  geom_point() +   # Values
  geom_point(aes(y = predict3), shape = 1) + # Fitted values
  theme_bw() 


```

```{r}
# sum of the squared residuals - linear model 
sum((ev_gg$residuals)^2)
# sum of the squared residuals - our guess of 3
sum((ev_gg$residual3)^2)
```

### A bonus lesson on the troubles of extrapolation

```{r}
# histogram of GDP growth
hist(ev$gdp_growth)

# test data frame
gdp_growth = ev$gdp_growth
elec_margin = ev$elec_margin
elec_margin[19] <- -4.4 # Trump's 2020 performance

#create new data frame
ps = tibble(gdp_growth, elec_margin)

# new model with 2020 
ev_fit2 = lm(elec_margin ~ gdp_growth, data = ps)

# new prediction
ps = ps %>% 
  mutate(predicted = predict(ev_fit),
         residuals = residuals(ev_fit))

#ggplot with new data and old regression line
ggplot(ps, aes(gdp_growth, elec_margin))+
  geom_point()+
  geom_abline(intercept = -.9292, slope = 3, color = "red")+
  ylim(-50,50)
  


ggplot(ps, aes(gdp_growth, elec_margin))+
  geom_smooth(method = "lm", se = FALSE, color = "lightgrey") +  # Plot regression slope
  geom_segment(aes(xend = gdp_growth, yend = predicted), alpha = .2) + #Segments show residual distance
  geom_label(aes(x= gdp_growth, y=(predicted+elec_margin)/2, label = round(residuals,3)), size = 3)  + # Labels show residual value                                                                  # alpha to fade lines
  geom_point() + # Values  
  geom_point(aes(y = predicted), shape = 1) + # Fitted values
  theme_bw()
```
