---
title: "lab_6_hypothesis_testing"
author: "Ben Burnley"
date: "2022-10-06"
output:
  pdf_document: default
  html_document: default
---

```{r}
# load packages 
library(tidyverse)
```

## Housekeeping for Problem Sets

-   Show work and code

-   If you work with a colleague, let me know

-   Clean and organized code

-   RMarkdown

    ```{r}
    # sample code
    data %>% 
      filter(x > 5) %>% 
      mutate(variable = x + y) %>% 
      summarize(mean = mean(variable))
    ```

## Confidence Intervals

Let's create a hypothetical population with approval for a president.

```{r}
set.seed(101)

# this simulation creates a population of 100,000 people asked 1 time whether they like they approve of the president
population = rbinom(100000, 1, .42)

# what is the population mean? 
mean(population)

# histogram of our data
hist(population)
```

### A poll of that population

```{r}
# a hypothetical poll of the 1500 members of the population
poll = sample(population, 1500, replace = FALSE)

# what is our sample estimate? 
x_bar = mean(poll)
x_bar
## standard deviation 
sd = sd(poll)
sd
# method for short cut of a standard deviation of a proportion if doing by hand
sqrt(x_bar*(1-x_bar))
```

The formula for calculating a confidence interval:

$$
C_{95} = \bar{X} \pm z_{C}SE_{\bar{X}} 
$$

Where, for this [example:\$\$\\bar{X}](example:$$\bar%7BX%7D){.uri} = .43\$\$

$$z_{C} = 1.96$$

$$SE_{\bar{X}} = \frac{.495}{\sqrt1500}
$$

```{r}
## we can figure this out! 

# estimate standard error
se = sd/sqrt(1500)
se  
#getting R to give us the critical value 
qt(.975, df = 1500)
z_c = qnorm(.975)
z_c
# calculate our margin 
margin = se * z_c
margin

# calculate interval 
upper_bound = x_bar + margin
lower_bound = x_bar - margin

print(c(lower_bound, upper_bound))
```

### Intuition about 95% Confidence Intervals

```{r}
## simulate the 100 samples 
sim_xbar = replicate(
  n = 100,
  mean(sample(population, 1500, replace = FALSE))
)
hist(sim_xbar) # check: what is this a simulation of?

# create data frame of 100 polls
polls_100 = data.frame(sim_xbar)

# add bounds and whether they contain our mean
polls_margins = polls_100 %>% 
  mutate(upper_bound = sim_xbar + margin,
         lower_bound = sim_xbar - margin,
         wegotit = factor(ifelse(lower_bound < .42, ifelse(upper_bound > .42, 1, 0), 0)),
         sample = seq(1:100))

## Check: what should the answer here be? 
sum(polls_margins$wegotit == 0)
```

### Plotting this exercise

```{r}
# set colors for graphing 
colorset = c('0'='red','1'='black')
## 
ggplot(polls_margins, aes(x = sample, y = sim_xbar)) + 
  geom_point(aes(color = wegotit)) + 
  geom_errorbar(aes(ymin = lower_bound, ymax = upper_bound, color = wegotit)) + 
  scale_color_manual(values = colorset) + 
  geom_hline(yintercept = .42, linetype = "dashed", color = "blue") + 
  labs(title = "95% Confidence Intervals") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  coord_flip()+
  theme_minimal()
  
```

## Difference in Means, in this Context

Are the estimates of two polls, taken from the same population, significantly different from one another.

**NOTE**: We should *not* expect this. Why?

What if instead we looked at two polls that sampled different regions?

Would we expect differences here?

```{r}
# population creations
population_north = rbinom(10000, 1, .54)
population_south = rbinom(10000, 1, .43)

#poll each region 
poll_north = sample(population_north, 150, replace = FALSE)
poll_south = sample(population_south, 75, replace = FALSE)

# estimate of approval in the north
mean_north = mean(poll_north)
mean_north

# estimate of approval in south
mean_south = mean(poll_south)
mean_south
```

Let's conduct a hypothesis test that the president's approval rating is the same in both regions.

What is $H_0$?

What is $H_A$?

```{r}
## we can test this!

# standard error for the north
se_north = sd(poll_north)/sqrt(150)
se_north

# standard error for the south
se_south = sd(poll_south)/sqrt(75)
se_south

# difference in means
diff_in_means = mean_north - mean_south
diff_in_means

# standard error of the difference
se_diff = sqrt(se_north^2 + se_south^2)
se_diff

t = (diff_in_means - 0)/se_diff
t 
```

stat_function(fun = dt, args = list(df = 1))+

```{r}
## can we reject the null? 
1 - pt(t, 74)

# one-tailed test 
tibble(x = seq(-3,3,0.01)) %>% 
ggplot(aes(x))+
  stat_function(fun = dt, args = list(df = 1))+
  geom_vline(aes(xintercept = 1.699), color = "green")+
  geom_vline(aes(xintercept = 1.665), linetype = "dashed")+
  geom_vline(aes(xintercept = 0), color = "blue")+
  theme_minimal()+
  labs(x = "t")+
    theme(
    axis.text.y = element_blank(),
    axis.title.y = element_blank()
  )
```

```{r}
tibble(x = seq(-3,3,0.01)) %>% 
ggplot(aes(x))+
  stat_function(fun = dt, args = list(df = 1))+
  geom_vline(aes(xintercept = 1.699), color = "red")+
  geom_vline(aes(xintercept = -1.99), linetype = "dashed")+
  geom_vline(aes(xintercept = 1.99), linetype = "dashed")+
  geom_vline(aes(xintercept = 0), color = "blue")+
  theme_minimal()+
  labs(x = "t")+
  theme(
    axis.text.y = element_blank(),
    axis.title.y = element_blank()
  )
```

## Hypothesis Testing with p-values

Say another poll from a hack pollster shows that the president's approval rating is at 50%. The media is calling asking you what you think of this. We think that is unlikely given our data, but can we do better than saying we disagree? Can we give $P(data|hypothesis)$?

What is $H_{0}$?

What is $H_{A}$?

```{r}
# remember calculations from earlier 
x_bar
se

(x_bar - .50)/se
pnorm()
```

### Test your own hypotheses

```{r}

```

## Sample Size and Hypothesis Testing

```{r}
plot_interval = function(n){
data = sample(population, n, replace = FALSE)
data_mean = mean(data)
data_sd = sd(data)
data_se = data_sd/sqrt(n)
z = qt(.975, df = n)
margin = z * data_se
tibble(seq(0,1,0.01)) %>% 
  ggplot(aes(x = 1, y = data_mean))+
  geom_point()+
  geom_errorbar(aes(ymin = data_mean - margin, ymax = data_mean + margin))+
  geom_hline(aes(yintercept = .42), linetype = "dashed", color = "blue")+
  theme_minimal()+
  coord_flip()+
  ylim(0,1)+
  theme(
    axis.text.y = element_blank(),
    axis.title.y = element_blank()
  )+
  labs(y = "Approval")}

plot_interval(10)
```
