---
title: 'lab_4_restructuring_data'
author: "Ben Burnley"
date: "2022-09-22"
output: html_document
---

## Reivew: ggplot2

```{r}
# packages 
library(ggplot2)
library(readr)
library(vtable)

# load data 
girlpower = read_csv("WomenGov10.csv")
extravars = read_csv("WomenGov10_extravars.csv")
View(girlpower)

# data vizualization
ggplot(girlpower, aes(citizen_ideology, num_women_total_hs, color = factor(termlimit)))+
  geom_point()+
  geom_smooth(method = "lm", se = F)+
  theme_classic()

vtable::sumtable(girlpower)

```

"

## 

## Introduction to the Tidyverse

```{r}
# load package 
#install.packages("tidyverse")
library(tidyverse)
```

The `tidyverse` package is really a family of packages that are super helpful for a number of data wrangling and visualizing needs. We will mostly be working with `dplyr` in this lesson, which allows us to restructure and reorganize our data.

One thing to note about the `tidyverse` is that it expects and really works best on data that is "tidy". Tidy data meets the following criteria:

1.  Every **column** is a **variable.**

2.  Every **row** is an **observation.**

3.  Every **cell** is a single **value.**

```{r}
# is this data tidy?
View(girlpower)
```

One final note: the `tidyverse` uses the pipe operator - `%>%` - in its syntax. The pipe functions by inheriting the data argument from the line above it. This allows the user to string together several verbs in a code chunk that is concise and has higher readability.

```{r}
# for example: this chunk of code creates two new variables
girlpower$regionid <- NA 
girlpower$regionid[girlpower$northeast==1] <- 1 
girlpower$regionid[girlpower$midwest==1] <- 2
girlpower$regionid[girlpower$west==1] <- 3
girlpower$regionid[girlpower$south==1] <- 4

girlpower$west <- NA
girlpower$west[girlpower$regionid==3] <- 1
girlpower$west[girlpower$regionid==1 |  #note the | (symbol for "or")
               girlpower$regionid==2 | 
               girlpower$regionid==4] <- 0


## same code in tidyverse
girlpower = girlpower %>% 
  mutate(regionid_tidy = ifelse(northeast == 1, 1,
                                 ifelse(midwest == 1, 2,
                                         ifelse(west == 1, 3,
                                                ifelse(south == 1, 4, NA)))),
         west_tidy =  ifelse(regionid == 3, 1, 0))
```

The base R version is 10 lines long and I have to type `girlpower` 16 times.

The tidy version is 5 lines long, I have to type `girlpower` once, AND there are no `$` .

### The `summarize` verb

```{r}
# put simply, you use this to summarize
girlpower %>% 
  summarize(mean_ideology = mean(pct_women_hs))

# lots of summary stats
girlpower %>% 
  summarize(mean = mean(pct_women_hs), 
            median = median(pct_women_hs),
            sd = sd(pct_women_hs),
            IQR = IQR(pct_women_hs),
            min = min(pct_women_hs),
            max = max(pct_women_hs))
```

This is not super useful on its own. Paired with other `tidyverse` verbs, it can be incredibly insightful.

### The `group_by` verb

This verb allows you to group your data by a variable of interest and then manipulate these groupings.

```{r}
# comprehension check: what has this done?
girlpower %>% 
  group_by(state)
```

```{r}
girlpower %>% 
  group_by(region) %>% 
  summarize(mean = mean(pct_women_hs))
```

### The `filter` verb

```{r}
girlpower %>% 
  filter(region %in% c("South", "Northeast"))
```

### The `mutate` verb

One of the most common tasks we will want to do is create new variables in our data set. The `mutate` verb does this, though the name is not necessarily intuitive - think "create".

```{r}
# the example above 
girlpower = girlpower %>% 
  mutate(regionid_tidy = ifelse(northeast == 1, 1,
                                 ifelse(midwest == 1, 2,
                                         ifelse(west == 1, 3, 4))),
         west_tidy =  ifelse(regionid == 3, 1, 0))

# creating a binary "conservative" variable 
girlpower = girlpower %>% 
  mutate(conservative_govt = govt_ideology < 50)

girlpower$conservative_govt

# new percent democratic variable 
girlpower = girlpower %>% 
  mutate(pct_dem = (num_dem_hs/num_total_hs)*100)

girlpower$pct_dem

# new LFP squared variable 
girlpower = girlpower %>% 
  mutate(pct_LFP_women_squared = pct_LFP_women^2)

girlpower$pct_LFP_women_squared
```

### `select` and `arrange`

```{r}
# selecting certain variables
girlpower %>% 
  select(state, conservative_govt)

# selecting a set of variables
girlpower %>% 
  select(state:termlimit)

# removing variables
girlpower %>% 
  select(-st)
```

```{r}
# arrange by lowest value
girlpower %>% 
  arrange(pct_women_hs)

# arange by highest value 
girlpower %>% 
  arrange(desc(pct_women_hs))
```

### Putting it All Together

These verbs vary in their use on their own, but put together they become very useful.

```{r}
# this is not super useful for a full data set, but comes in handy when summarizing
girlpower %>% 
  group_by(region) %>% 
  summarize(mean = mean(pct_women_hs)) %>% 
  arrange(desc(mean))
```

```{r}
# grouped mutate example - with ungroup()
girlpower %>% 
  group_by(region) %>% 
  mutate(avg_region_pct_women = mean(pct_women_hs)) %>% 
  ungroup() %>% 
  mutate(above_region_avg = ifelse(pct_women_hs > avg_region_pct_women, 1, 0)) %>% 
  select(state, pct_women_hs, avg_region_pct_women, above_region_avg)
```

## Merges

Another extremely helpful tool available from the tidyverse is are the joining verbs that allow you to merge together different data sets.

### Mutating Joins

```{r}
view(extravars)

# leftjoin - match rows from b to a 
left = girlpower %>% 
  left_join(extravars, by = "state")

# right join - match rows from a to b 
right = girlpower %>% 
  right_join(extravars, by = "state")

# inner join - join and keep only matches 
inner = girlpower %>% 
  inner_join(extravars, by = "state")

# full join - keep all values
full = girlpower %>% 
  full_join(extravars, by = "state")
```

### Filtering Joins

```{r}
# semi join - all the rows in a that have a match in b
semi = girlpower %>% 
  semi_join(extravars, by = "state")

# antijoin - all rows in a that DO NOT have match in b
anti = girlpower %>% 
  anti_join(extravars, by = "state")
```
