---
title: "lab_12_interactions"
author: "Ben Burnley"
date: "2022-12-01"
output: html_document
---

```{r}
# libraries 
library(tidyverse)
#intstall.packages("corrplot")
library(corrplot)
#install.packages("dotwhisker")
library(dotwhisker)
library(stargazer)
#install.packages("moderndive")
library(moderndive)

# load data 
data = read_csv("speeding_tickets_text.csv")
```

Codebook

MPHOver - Miles per hour over the speed limit

Amount - Assessed fine for the ticket (in dollars)

Age - Age of driver (in years)

Female - Equals 1 for women and 0 for men

Black - Equals 1 if driver was Black and 0 otherwise

Hispanic - Equals 1 if driver was Hispanic and 0 otherwise

StatePol - Equals 1 if ticketing officer was state patrol officer

OutTown - Equals 1 if driver was from out of town and 0 otherwise

OutState - Equals 1 if driver was from out of state and 0 otherwise

### Simple Interaction Model

Consider the model:

$$\hat{y} = \beta_0 + \beta_1Female + \beta_2OutState + \beta_3FemaleXOutState$$

```{r, warning=F, message=FALSE}
# run this model in an lm function
m1 = lm(Amount ~ Female + OutState + Female*OutState, data = data)

# model summary
summary(m1)

# plot of this model
ggplot(data, aes(Female, Amount, color = factor(OutState)))+
  geom_point(alpha = 0.5)+
  moderndive::geom_parallel_slopes()+
  scale_color_manual(values = c("darkred","darkblue"), name = "Out of State")+
  theme_minimal()
```

### Another Example

Consider the model:

$$\hat{y} = \beta_0 + \beta_1Female + \beta_2Black + \beta_3FemaleXBlack$$

```{r,message=F, warning=F}
# run this model in an lm function
m2 = lm(Amount ~ Female + Black + Female*Black, data = data)

# model summary
summary(m2)

# plot of this model
ggplot(data, aes(Female, Amount, color = factor(Black)))+
  geom_point(alpha = 0.5)+
  moderndive::geom_parallel_slopes()+
  scale_color_manual(values = c("darkred","darkblue"), name = "Black")+
  theme_minimal()
```

### With a continuous IV of Interest

Consider the model:

$$\hat{y} = \beta_0 + \beta_1MPHOver + \beta_2Female + \beta_3MPHOverXFemale$$

```{r, message=F,warning=FALSE}
m3 = lm(Amount ~ MPHover + Female + MPHover*Female, data = data)

# model summary
summary(m3)

# plot of this model
ggplot(data, aes(MPHover, Amount, color = factor(Female)))+
  geom_point(alpha = 0.5)+
  geom_smooth(method = "lm", se = F)+
  scale_color_manual(values = c("darkred","darkblue"), name = "Female")+
  theme_minimal()
```
