---
title: "VDEM Presentation"
author: "Ben Burnley"
date: "2022-11-01"
output: html_document
---

Outline
 - Definition and Conceptualization 
 - Explore the distribution 
 - Relate 3-4 additional variables that you expect to prove related to your variables. Write your hypotheses before measurement. 
 - Explore the hypotheses empirically through data visualization or statistical testing. Did you hypotheses pan out? Why or why not? 
 - 15 - 20 mins
 
 Variables: 
 Internet binary (C) (v2mecenefibin)
 Online media existence (C) (v2smonex)
 Online media perspectives (C) (v2smonper)
 Online media fractionalization (C) (v2smmefra)
 Elites’ use of social media to organize offline action (C) (v2smorgelitact)
 Types of organization through social media (C) (v2smorgtypes)
 Other types of organization through social media (C) (v2smorgtypestxt)
 Party/candidate use of social media in campaigns (C) (v2smcamp)
 
 
 
 
```{r}
# libraries 
library(tidyverse)
library(vdemdata)
library(ggridges)
library(maps)
library(MetBrewer)

install.packages("devtools")
# now, install the vdemdata package directly from GitHub
devtools::install_github("vdeminstitute/vdemdata")

# import 
data = vdemdata::vdem
world_map = map_data("world")


my_theme =  theme_minimal()+
  theme(
    plot.title = element_text(face = "bold", size = 16)
  )
```

### Cleaning 
```{r}
first = data %>% 
  filter(year > 1992) %>% 
  select(country_name) %>% 
  distinct(country_name)

second = world_map %>% 
  select(region) %>% 
  distinct(region)

first %>% 
  anti_join(second, by = c("country_name" = "region"))


clean = data %>% 
  mutate(country_name = if_else(country_name == "Burma/Myanmar", "Myanmar",
                                ifelse(country_name == "United States of America", "USA",
                                       ifelse(country_name == "United Kingdom", "UK",
                                              ifelse(country_name =="Republic of the Congo", "Republic of Congo",
                                                            ifelse(country_name == "The Gambia", "Gambia", country_name)))))) %>% 
  mutate(censor = 3 - v2mecenefi_osp)
```

 
```{r, message=FALSE}
# eda for v2smonex
summary(data$v2smonex_osp)

ggplot(data, aes(v2smonex_osp))+
  geom_histogram()

data %>% 
  filter(year > 2000) %>% 
  group_by(year) %>% 
  ggplot(aes(v2smonex_osp, factor(year), group = year, fill = factor(stat(quantile))))+
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE,
    quantiles = 4, quantile_lines = TRUE, alpha = .75)+
  scale_fill_viridis_d(name = "Quartiles", guide = FALSE)+
  labs()
  theme_ridges()

```
 
 ## Internet Binary 
```{r}
## summary of the variable
summary(data$v2mecenefibin)
table(data$v2mecenefibin_osp)

# increase in internet over time
data %>% 
  filter(year > 1992) %>% 
  group_by(year) %>% 
  summarize(mean_int = mean(v2mecenefibin_osp, na.rm = T)) %>% 
  ggplot(aes(year, mean_int))+
  geom_line(color = "blue", size = 2)+
  labs(
    title = "Growth of Internet Access over time",
    caption = "Source: Varieties of Democracy Project",
    x = "Year",
    y = "% of Countries with Internet Access"
  )+
  theme_minimal()+
  theme(
    plot.title = element_text(face = "bold", size = 16)
  )+
  my_theme

## mapping the adoption of internet 
clean %>% 
  filter(year %in% c(1993,1998,2003,2008,2013,2018)) %>% 
  select(country_name, year, v2mecenefibin_osp) %>% 
  ggplot(aes(fill = v2mecenefibin_osp))+
  geom_map(aes(map_id = country_name), map = world_map)+
  scale_fill_viridis_c(option = "D", guide = "none")+
  expand_limits(x = world_map$long, y = world_map$lat)+
  facet_wrap(~year)+
  theme_void()+
  theme(
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    strip.text = element_text(face = "bold"),
    plot.title = element_text(face = "bold"),
    plot.background = element_rect(fill = "white", color = "white"),
    text = element_text(color = "black")
    )+
  labs(
    title = "Internet Access over time",
    subtitle = "Lighter colors indicate more access",
    caption = "Source: Varities of Democracy Project"
  )

```
### 
```{r}
clean %>% 
  select(country_name, year, v2smonex_osp)%>% 
  filter(year %in% c(2000, 2005, 2010, 2015, 2020)) %>% 
  ggplot(aes(v2smonex_osp, factor(year), group = year, fill = factor(stat(quantile))))+
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE,
    quantiles = 4, quantile_lines = TRUE, alpha = .75)+
  scale_fill_viridis_d(name = "Quartiles", guide = "none", alpha = 0.75)+
  labs(
    title = "Online Media Existence over Time",
    x = "Online Media Existence",
    y = "Year",
    caption = "Source: Varieties of Democracy Project"
      )+
  theme_ridges()+
  theme()

clean %>% 
  filter(year %in% c(2000, 2005, 2010, 2015, 2020)) %>% 
  select(country_name, year, v2smonex_osp) %>% 
  ggplot(aes(fill = factor(round(v2smonex_osp))))+
  geom_map(aes(map_id = country_name), map = world_map)+
  scale_fill_viridis_d(name = "OME Score")+
  expand_limits(x = world_map$long, y = world_map$lat)+
  facet_wrap(~year)+
  theme_void()+
  theme(
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    strip.text = element_text(face = "bold"),
    plot.title = element_text(face = "bold", hjust = .5),
    plot.background = element_rect(fill = "white", color = "white"),
    text = element_text(color = "black"),
    legend.position = c(.85,.3),
    )+
  labs(
    title = "Online Media Existence",
    caption = "Source: Varieties of Democracy Project"
  )

clean %>% 
  filter(year > 2000) %>% 
  group_by(year) %>% 
  summarize(mean_int = mean(v2smonex_osp, na.rm = T)) %>% 
  ggplot(aes(year, mean_int))+
  geom_line(color = "blue", size = 2)+
  labs(
    title = "Growth of Online Media Existence",
    caption = "Source: Varieties of Democracy Project",
    x = "Year",
    y = "Average Score"
  )+
  
  theme_minimal()+
  theme(
    plot.title = element_text(face = "bold", size = 16)
  )+
  my_theme
```
## Party Social Media Use Online
```{r}
v2smcamp_osp

clean %>% 
  filter(year > 1999) %>% 
  group_by(year) %>% 
  summarize(mean_int = mean(v2smcamp_osp, na.rm = T)) %>% 
  ggplot(aes(year, mean_int))+
  geom_line(color = "blue", size = 2)+
  labs(
    title = "Growth of Social Media Use by Party/Candidate",
    caption = "Source: Varieties of Democracy Project",
    x = "Year",
    y = "Average Score"
  )+
  
  theme_minimal()+
  theme(
    plot.title = element_text(face = "bold", size = 16)
  )+
  my_theme

clean %>% 
  select(country_name, year, v2smcamp_osp)%>% 
  filter(year %in% c(2000, 2005, 2010, 2015, 2020)) %>% 
  ggplot(aes(v2smcamp_osp, factor(year), group = year, fill = factor(stat(quantile))))+
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE,
    quantiles = 4, quantile_lines = TRUE, alpha = .75)+
  scale_fill_viridis_d(name = "Quartiles", guide = "none", alpha = 0.75)+
  labs(
    title = "Social Media Use by Party/Candidate over Time",
    x = "Social Media Usage Score",
    y = "Year",
    caption = "Source: Varieties of Democracy Project"
      )+
  theme_ridges()+
  theme()


clean %>% 
  filter(year %in% c(2000, 2005, 2010, 2015, 2020)) %>% 
  select(country_name, year, v2smcamp_osp) %>% 
  ggplot(aes(fill = factor(round(v2smcamp_osp))))+
  geom_map(aes(map_id = country_name), map = world_map)+
  scale_fill_viridis_d(name = "Social Media Usage")+
  expand_limits(x = world_map$long, y = world_map$lat)+
  facet_wrap(~year)+
  theme_void()+
  theme(
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    strip.text = element_text(face = "bold"),
    plot.title = element_text(face = "bold", hjust = .5),
    plot.background = element_rect(fill = "white", color = "white"),
    text = element_text(color = "black"),
    legend.position = c(.85,.3),
    )+
  labs(
    title = "Social Media Use by Party/Candidate over Time",
    caption = "Source: Varieties of Democracy Project"
  )
```

```{r}
v2mecenefi_osp

clean %>% 
  filter(year > 1993) %>% 
  group_by(year) %>% 
  summarize(mean_int = mean(censor, na.rm = T)) %>% 
  ggplot(aes(year, mean_int))+
  geom_line(color = "blue", size = 2)+
  labs(
    title = "Internet Censorship Effort",
    caption = "Source: Varieties of Democracy Project",
    x = "Year",
    y = "Average Score"
  )+
  theme_minimal()+
  theme(
    plot.title = element_text(face = "bold", size = 16)
  )+
  my_theme

clean %>% 
  select(country_name, year, censor)%>% 
  filter(year %in% c(1995, 2000, 2005, 2010, 2015, 2020)) %>% 
  ggplot(aes(censor, factor(year), group = year, fill = factor(stat(quantile))))+
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE,
    quantiles = 4, quantile_lines = TRUE, alpha = .75)+
  scale_fill_viridis_d(name = "Quartiles", guide = "none", alpha = 0.75)+
  labs(
    title = "Internet Censorship Effort over Time",
    x = "Internet Censorship Score",
    y = "Year",
    caption = "Source: Varieties of Democracy Project"
      )+
  theme_ridges()+
  theme()


clean %>% 
  filter(year %in% c(1995, 2000, 2005, 2010, 2015, 2020)) %>% 
  select(country_name, year, censor) %>% 
  ggplot(aes(fill = factor(round(censor))))+
  geom_map(aes(map_id = country_name), map = world_map)+
  scale_fill_viridis_d(name = "Internet Censorship")+
  expand_limits(x = world_map$long, y = world_map$lat)+
  facet_wrap(~year)+
  theme_void()+
  theme(
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    strip.text = element_text(face = "bold"),
    plot.title = element_text(face = "bold", hjust = .5),
    plot.background = element_rect(fill = "white", color = "white"),
    text = element_text(color = "black"),
    legend.position = "bottom",
    )+
  labs(
    title = "Internet Censorship Effort over Time",
    caption = "Source: Varieties of Democracy Project"
  )
```

```{r,message=F,warning=F}
## internet vs. online media adoption

d1 = clean %>% 
  filter(year > 1992 & year < 2000) %>% 
  select(country_name, year, v2mecenefibin_osp)

summary(lm(v2mecenefibin_osp ~ year, d1))



ggplot(clean, aes(v2mecenefibin_osp,v2smonex_osp))+
  geom_point()+
  geom_smooth(method = "lm")

summary(glm(v2mecenefibin_osp ~ v2smonex_osp, family = "binomial", data = clean))

summary(lm(censor ~ v2smonex_osp, data = clean))

ggplot(clean, aes(v2smonex_osp,censor))+
  geom_point(position = "jitter", alpha = 0.5)+
  geom_smooth(method = "lm", se = F)+
  labs(
    title = "Relationship between Online Media and Censorship Effort",
    x = "Online Media Existence",
    y = "Internet Censorship Effort",
    caption = "Source: Varieties of Democracy Project"
  )+
  my_theme

## internet adoption overtime 
clean %>% 
  filter(year > 1992 & year < 2000) %>% 
  ggplot(aes(year, v2mecenefibin_osp))+
  geom_point(alpha = .75, position = "jitter")+
  geom_smooth(method = "lm", se = F)+
  labs(
    title = "Internet Adoption over time",
    caption = "Source: Varieties of Democracy Project",
    x = "Year",
    y = "Likelihood of Internet Adoption"
  )+
  my_theme

summary(lm(v2mecenefibin_osp ~ year, data = d1))
summary(lm(censor ~ v2smonex_osp, data = clean))

```
 
 
```{r}
clean %>% 
  ggplot(aes(v2smonex_osp, v2smcamp_osp))+
  geom_point(alpha = .75)+
  geom_smooth(method = "lm")+
  labs(
    title = "Relationship between Online Media and Party/Candidate Usage",
    x = "Online Media Existence",
    y = "Party/Candidate Usage of Social Media",
    caption = "Source: Varieties of Democracy Project"
  )+
  my_theme

summary(lm(v2smcamp_osp ~ v2smonex_osp, data = clean))
```

```{r}
clean %>% 
  ggplot(aes(v2smcamp_osp,censor))+
  geom_point(alpha = .5, position = "jitter")+
  geom_smooth(se = F)+
  labs(
    title = "Relationship between Party/Candidate Usage & Censorship",
    caption = "Source: Varieties of Democracy Project",
    x = "Party/Candidate Social Media Usage",
    y = "Internet Censorship Effort"
  )+
  my_theme

summary(lm(censor ~ v2smcamp_osp, data = clean))
```
 
```{r}
clean %>%
  filter(year > 1999) %>% 
  ggplot(aes(v2smcamp_osp,censor, color = year))+
  geom_point(alpha = .5, position = "jitter")+
  geom_smooth(method = "lm", se = F)+
  scale_color_viridis_c(name = "Year")+
  labs(
    title = "Relationship between Party/Candidate Usage & Censorship",
    caption = "Source: Varieties of Democracy Project",
    x = "Party/Candidate Social Media Usage",
    y = "Internet Censorship Effort"
  )+
  my_theme
```
```{r}
clean %>%
  filter(year > 2010) %>% 
  ggplot(aes(v2smcamp_osp,censor, color = v2smonex_osp))+
  geom_point(alpha = .5, position = "jitter")+
  geom_smooth(method = "lm", se = F)+
  scale_color_viridis_c(name = "Online Media")+
  labs(
    title = "Relationship between Party/Candidate Usage & Censorship",
    caption = "Source: Varieties of Democracy Project",
    x = "Party/Candidate Social Media Usage",
    y = "Internet Censorship Effort"
  )+
  my_theme
```
 
```{r}
c1 = clean %>% 
  filter(country_text_id %in% c("SAU", "CHN", "ARE"))

clean %>%
  filter(year > 1999) %>% 
  ggplot(aes(v2smcamp_osp,censor, color = v2smonex_osp))+
  geom_point(alpha = .2, position = "jitter")+
  geom_text(aes(label = country_text_id), alpha = 1, color = "black")+
  geom_smooth(method = "lm", se = F)+
  scale_color_viridis_c(name = "Online Media")+
  labs(
    title = "Relationship between Party/Candidate Usage & Censorship",
    caption = "Source: Varieties of Democracy Project",
    x = "Party/Candidate Social Media Usage",
    y = "Internet Censorship Effort"
  )+
  my_theme


```
 
 